Modal Verification Patterns for Systems

### 2025-06-23 paper.tex
+

Filesystems do not only handle the address-translation from a disk-page to a disk-block but also need to handle the different \emph{modes} of computing in which addressing disk resources occur at different consistency modes: a disk block persisted on the disk and a disk-page which may not be persisted on the disk yet. To handle the consistency of the disk resources in the existence of different modes of execution, filesystems employ different policies when indexing the disk resource. Which data-structure is going to be used to implement \emph{mappi
\textsf{ModFS} is a copy-on-write (\textsf{CoW}) filesystem, which as described in the introduction performs updates by copying modified data to \emph{new and unused} locations on disk rather than updating in place, ensuring that in the absence of physical storage corruption, there is \emph{always} a fully consistent filesystem on disk, removing the need for recovery from journals in the case of a crash, which can be very expensive for large volumes.
%\textsf{ModFS} is implemented in \textsc{Go}~\cite{donovan2015go}, using the FUSE~\cite{fuse} interface to run a filesystem in usermode. \textsf{ModFS}'s code is translated to \textsf{Roqc} using the \textsf{Goose}~\cite{chajed2019verifying,chajed2021gojournal} tool, which translates a significant subset of \textsf{Go} to \textsf{Roqc} to permit formal verification of \textsc{Go} code using \textsf{Iris}~\cite{jung2018iris}. \textsf{Goose} generates code in an extension of \textsf{Iris}'s \textsf{HeapLang} with libraries for disk manipulation and \textsc{Go} primitive types (e.g., maps) as shown in Figure \ref{fig:modfslayout}.

@@

%\textsf{ModFs} is a \textsf{CoW} file system where the mapping of the logical locations, e.g., file pages indexed by file offsets, to disk locations by employing a \textsf{CoW} tree \cite{Astrahan1976SystemR,shadowclone} structure. In Figure \ref{fig:cow_index_tree}, we see that accessing and updating the $\textsf{page}_n$ of file \textsf{"foo"} referred by the file offset $\textsf{offset}$ 
%\[ \textsf{write}(\textit{foo},\textsf{offset})\]
%requires locating the corresponding physical disk block of the in-memory $\textsf{page}_n$ whose mapping is establised the \textit{copy-on-write} index tree (in the middle layer). Like Btrfs ~\cite{}, \textsf{ModFs} indexes all filesystem objects (e.g. files and directories) as a nested trees, and to map a logical page to a physical page, the file system first locates the root of the nested index-tree belonging to \textsf{"foo"}.
Semantically, in \textsf{ModFS} (as in other \textsf{CoW} filesystems~\cite{young1992episode,bonwick2003zettabyte,rodeh2013btrfs,Rosenblum1992LFS,hitz1994file}) any update to file data or metadata is realized on  unused disk locations by treating the tree as a persistent data structure. 
As an example, Figure \ref{fig:cow_index} shows that an update to the file $\textsf{"foo"}$ at $\textsf{"offset"}$ ends up with locating the related disk chunk (a set of blocks) in the gray region, copying it to newly allocated disk chunks (dotted chunks on the right side of the bold long-dash-point line of the middle layer withing dotted section), and performing the update on these newly-allocated disk blocks. This requires further updates to those metadata blocks which previously referred to the old block locations to instead reference the newly-allocated blocks. This must be repeated until the allocation/update of the new chunks percolates up to the root (including the root), through a process called \textsf{\textit{write-amplification}}.

@@

After this process, traversing the filesystem from either the old or new root yields consistent views of the filesystem at different points in time, each commonly referred to as a \emph{snapshot}.
These snapshots overlap on the unmodified portions of the filesystem (the left-most grid section), with the reclaimed the chunks (the middle gray section) becoming available for reuse once the new snapshot becomes the official current state of the filesystem.\footnote{Many of these filesystems permit keeping archived snapshots of the filesystem state, e.g., a snapshot of the system every hour up to a certain number of days, to support user data recovery. In these cases only disk regions that are not still in use by an archived snapshot are reclaimed.}. This permits the filesystem to atomically switch between consistent on-disk states, by updating the oldest so-called \emph{super} chunk at a fixed location on disk.\footnote{Filesystems must take care to ensure that possibly-flaky hardware truly does perform an atomic update.} Super chunks, as we see in Figure \ref{fig:cow_index} as ($\textsf{Super}_0$ and $\textsf{Super}_1$), are the first two blocks of the disk dedicated to keep the state of the file system which is composed of some meta-data on the filesystem, and most importantly to the root chunk of the file system index-tree. Therefore, in case of a crash during the update to the \textsf{file "foo"}, we can reboot the file system from the previous persisted root, \textsf{root chunk}, via reading the $\textsf{Super}_0$. This means the whole disk consistency checks are unnecessary when recovering from a crash~\cite{bonwick2003zettabyte,hitz1994file}; recovery is immediate, as opposed to requiring hours or days~\cite{henson2006reducing,agrawal2009speedy} (or longer~\cite{henson2006chunkfs}) for recovering a large volume’s consistency from logs. 
An important additional subtlety with this kind of filesystem is \emph{when} blocks in the reclaimed region can actually be reused: only \emph{after} the snapshot that made those blocks ``free'' has been completely written to disk. Allowing reuse before this could result in in-place updates: part of the new snapshot could attempt to store data in space used by the prior snapshot, allowing inconsistent on-disk structures. 
\begin{figure}
    \includegraphics[width=\columnwidth]{journalling.png}
    \caption{A Filesystem Utilizing \emph{Journalling} for Consistent Disk-Layout ~\cite{chajed2021gojournal}}
    \label{fig:journal}
\end{figure}

>>>>

Filesystems do not only handle the address-translation from a disk-page to a disk-block but also need to handle the different \emph{modes} of computing in which addressing disk resources occur at different consistency modes: a disk block persisted on the disk and a disk-page which may not be persisted on the disk yet. To handle the consistency of the disk resources in the existence of different modes of execution, filesystems employ different policies when indexing the disk resource. Which data-structure is going to be used to implement address-traslation from disk-pages to disk-blocks? What is going to be the \emph{update-policy} to the indexing map implementing address-translation which defines the \emph{consistency-policy} for the disk resources? Different filesystem implementations give different answers to these questions. One example answer to updating policy is copy-on-write (\textsf{CoW}). A (\textsf{CoW}) filesystem performs updates by copying modified data to \emph{new and unused} locations on disk rather than updating \emph{in-place}, ensuring that in the absence of physical storage corruption, there is \emph{always} a fully consistent filesystem on disk, removing the need for recovery from journals in the case of a crash, which can be very expensive for large volumes.
%\textsf{ModFS} is implemented in \textsc{Go}~\cite{donovan2015go}, using the FUSE~\cite{fuse} interface to run a filesystem in usermode. \textsf{ModFS}'s code is translated to \textsf{Roqc} using the \textsf{Goose}~\cite{chajed2019verifying,chajed2021gojournal} tool, which translates a significant subset of \textsf{Go} to \textsf{Roqc} to permit formal verification of \textsc{Go} code using \textsf{Iris}~\cite{jung2018iris}. \textsf{Goose} generates code in an extension of \textsf{Iris}'s \textsf{HeapLang} with libraries for disk manipulation and \textsc{Go} primitive types (e.g., maps) as shown in Figure \ref{fig:modfslayout}.

@@

%\textsf{ModFs} is a \textsf{CoW} file system where the mapping of the logical locations, e.g., file pages indexed by file offsets, to disk locations by employing a \textsf{CoW} tree \cite{Astrahan1976SystemR,shadowclone} structure. In Figure \ref{fig:cow_index_tree}, we see that accessing and updating the $\textsf{page}_n$ of file \textsf{"foo"} referred by the file offset $\textsf{offset}$ 
%\[ \textsf{write}(\textit{foo},\textsf{offset})\]
%requires locating the corresponding physical disk block of the in-memory $\textsf{page}_n$ whose mapping is establised the \textit{copy-on-write} index tree (in the middle layer). Like Btrfs ~\cite{}, \textsf{ModFs} indexes all filesystem objects (e.g. files and directories) as a nested trees, and to map a logical page to a physical page, the file system first locates the root of the nested index-tree belonging to \textsf{"foo"}.
Semantically, in \textsf{ModFS}~\cite{} (as in other \textsf{CoW} filesystems~\cite{young1992episode,bonwick2003zettabyte,rodeh2013btrfs,Rosenblum1992LFS,hitz1994file}) any update to file data or metadata is realized on  unused disk locations by treating the tree as a persistent data structure. 
As an example, Figure \ref{fig:cowindex} shows that an update to the file $\textsf{"foo"}$ at $\textsf{"offset"}$ ends up with locating the related disk chunk(multiple disk-pages) in the gray region, copying it to newly allocated disk chunks (dotted chunks on the right side of the bold long-dash-point line of the middle layer withing dotted section), and performing the update on these newly-allocated disk blocks. This requires further updates to those metadata blocks which previously referred to the old block locations to instead reference the newly-allocated blocks. This must be repeated until the allocation/update of the new chunks percolates up to the root (including the root), through a process called \textsf{\textit{write-amplification}}.

@@

Like VMMs navigating between different address-spaces, a \textsf{CoW} filesystem switches from either the old to the new root yields consistent views of the filesystem at different points in time, each commonly referred to as a \emph{snapshot}.
These snapshots overlap on the unmodified portions of the filesystem (the left-most grid section), with the reclaimed the chunks (the middle gray section) becoming available for reuse once the new snapshot becomes the official current state of the filesystem.\footnote{Many of these filesystems permit keeping archived snapshots of the filesystem state, e.g., a snapshot of the system every hour up to a certain number of days, to support user data recovery. In these cases only disk regions that are not still in use by an archived snapshot are reclaimed.}. This permits the filesystem to atomically switch between consistent on-disk states, by updating the oldest so-called \emph{super} chunk at a fixed location on disk.\footnote{Filesystems must take care to ensure that possibly-flaky hardware truly does perform an atomic update.} Super chunks, as we see in Figure \ref{fig:cowindex} as ($\textsf{Super}_0$ and $\textsf{Super}_1$), are the first two blocks of the disk dedicated to keep the state of the file system which is composed of some meta-data on the filesystem, and most importantly to the root chunk of the file system index-tree. Therefore, in case of a crash during the update to the \textsf{file "foo"}, we can reboot the file system from the previous persisted root, \textsf{root chunk}, via reading the $\textsf{Super}_0$. This means the whole disk consistency checks are unnecessary when recovering from a crash~\cite{bonwick2003zettabyte,hitz1994file}; recovery is immediate, as opposed to requiring hours or days~\cite{henson2006reducing,agrawal2009speedy} (or longer~\cite{henson2006chunkfs}) for recovering a large volume’s consistency from logs. 
An important additional subtlety with this kind of filesystem is \emph{when} blocks in the reclaimed region can actually be reused: only \emph{after} the snapshot that made those blocks ``free'' has been completely written to disk. Allowing reuse before this could result in in-place updates: part of the new snapshot could attempt to store data in space used by the prior snapshot, allowing inconsistent on-disk structures. 
\begin{figure}
    \includegraphics[width=\columnwidth]{journalling.png}
    \caption{A Filesystem Utilizing \emph{Journalling} for Consistent Disk-Layout ~\cite{chajed2021gojournal}}
    \label{fig:journal}
\end{figure}
